Improving the Replicability of Psychological Science Through Pedagogy

### 2017-07-02 paper/paper.Rmd
mcf tweaks

We begin by describing the details of our course projects to give a sense of what a classroom replication project entails. We then discuss broader decision points for the course design that other instructors might consider. Lastly, we report the results our replications, highlighting the importance of wide-scale replication efforts.

@@

\specialcell{Citation\\(Psychological Science; 2015)} & Expt. & \specialcell{Open\\Data?} & \specialcell{Open\\Materials?} & \specialcell{On\\MTurk?} & \specialcell{N\\(orig)} & \specialcell{Power\\Standard} & \specialcell{N\\(rep)} & \specialcell{Instructor\\Fidelity}\\ 
  \hline
Atir, Rosenzweig, \& Dunning & 1b & No & Yes & Yes & 202 & Other &  50 & 6.67 & \\ 
Ko, Sadler, \& Galinksy & 2 & Yes & Some & No &  40 & Original &  40 & 4.67 &\\ 
Lewis \& Oyserman & 4 & No & Yes & Yes & 122 & 80\% power & 128 & 6.67 &\\ 
Liverence \& Scholl & 1 & No & Some & No &  18 & Original* &  19 & 5.33 &\\ 
Proudfoot, Kay, \& Koval & 1 & No & No & Yes &  80 & 80\% power &  84 & 6.67&\\ 
Scopelliti, Loewenstein, \& Vosgerau & 3 & Yes & Yes & Yes & 550 & Other & 124 & 5.67& \\ 
Sofer et al. & 1 & Yes & Yes & No & 48 & Other &  95 & 4.33& \\ 
Storm \& Stone & 3 & No & No & No & 48 & Original &  61 & 4.00& \\ 
Wang et al. & 2 & No & No & No & 219 & 80\% power & 397 & 4.33&\\ 
Xu \& Franconeri & 1a & No & No & No &  12 & Other* &  27 & 5.00& \\ 
Zaval, Markowitz, \& Weber & 1 & Yes & Yes & Yes & 312 & Original & 321 & 5.00&\\ 
   \hline
\end{tabular}
\caption{\label{tab:summary} Summary characteristics of original studies and our replications. All project materials available publicly at osf.io/98ta4/files/, and links to individual project preregistrations, reports, and web experiments are available at osf.io/98ta4/wiki/ * marks projects where the number of trials was modified.} 
\label{tab:summary}
\end{table}

@@

RXDH, ENS, and MCF designed the project, supported data planning and data collection for all projects, analyzed the data, and wrote the paper. CA, JMA, RC, EH, MK, AL, SR, JR, SS, JS, and JT planned their individual projects, programmed their studies, collected their data, and analyzed their data.

@@

Though others' courses that incorporate replication projects will look quite different from our own, we recommend having the replication as a centerpiece of the course and giving ample opportunity for feedback throughout the study design. Without a central focus on the replication project, students would not have the time or motivation to complete a high quality project; without multiple opportunities for instructor feedback, we would have far less confidence in the quality of the final projects.

@@

In our class, students selected projects based on interest, rather than via a systematic sampling strategy. After selecting a paper, students chose a particular study from within that paper based on their judgment of interest and feasibility. We encouraged students with little programming background to opt for methodologically simple studies, and more advanced students to take on a challenging design. This level of freedom is not ideal for less advanced courses, and the set of available replication studies should likely be curated for students at other levels.  For example, in a large undergraduate class it may be logistically simpler for the instructors to choose 1 -- 3 replication projects that students can select from. This step would both give instructors a greater degree of familiarity with the literature and allow for the development of more scaffolding with respect to the development of experimental materials. A mid-level course might split the difference between these two models, providing students with a broader -- but still curated -- list of potential projects.

@@

Some of these tools can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, typically have relatively little experience with web-based empirical studies or literate, reproducible data analysis. Yet by the end of the course, all gain sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. Our experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environment -- students can quickly learn to use a reproducible toolset.

@@

Determining appropriate sample sizes is a major challenge in replication research [@button2013;@simonsohn2015]. Sample planning based on analysis of the achieved effect size in a previously-published study is problematic because of the likelihood of inflation of published effect sizes due to the "winner's curse" [@hoenig2001;@button2013]. But -- especially within the constraints of a limited budget -- it can be impossible to follow more conservative guidelines in all cases. For example, @simonsohn2015 recommends 2.5x the original sample, which can be feasible for small or under-powered original studies but may lead to impractical or unnecessary recommendations for studies that were initially large and/or adequately powered. In general, samples for course projects will be limited by the available resources. But for all replication research of the type advocated here, achieving high power for a smaller number of studies, perhaps conducted by groups, is preferable to conducting more lower-power studies, because results are much more likely to make a contribution to the replication literature (as well as to be interpretable by students). Regardless of the eventual sample collected, we find that the discussion of sample size determination is one of the most eye-opening parts of our class and so we encourage instructors to allow students to participate in the planning process.

@@

Over half of our students contacted original authors to request study materials or clarification on methods and/or analysis, but for the purposes of our class, we chose not to require contact with all original authors. We strongly believe methods and materials should be open to allow for replication and productive science [@nosek2015], and hoped to empower students to engage with and build on the published literature directly -- without personal relationships or even personal contact. That is, we chose to do good-faith replications given all information in the original articles and supplemental online materials. Authors were contacted only in cases where there were issues about access to materials or ambiguities in design or analysis. There are benefits of contacting original authors, however. In addition to professional courtesy, authors can provide valuable feedback or guidance (albeit at the cost of some imposition) prior to data collection, potentially heading off post-hoc debates about methodological choices. In the end, it is up to individual instructors to decide what they believe is best for their students.

@@

Institutional Review Boards (IRBs) are idiosyncratic across institutions, but we suspect certain ethical concerns will commonly arise when attempting to implement class replication projects. First, instructors may first be concerned that IRB approval is not possible within the time constraints of the class.  To mitigate this issue, we suggest contacting the IRB chair well in advance of the class to determine whether to create a standard protocol that encompasses all replications that will be conducted, or to determine a timeline for submitting individual protocols very early in the course. Both issues may be simplified if the study or pool of studies to be replicated is pre-selected by instructors. Second, review boards may be concerned about risks particular to the collection of data by students, e.g., lack of debriefing or greater risk of breach of confidentiality due to novice experimenters. We recommend building training around ethical issues into the syllabus of replication-based classes. This practice is both positive for students and mitigates potential IRB concerns via explicit training.

>>>>

We begin by providing the details of our course projects to give a sense of what a classroom replication project entails. We then discuss broader decision points for the course design that other instructors might consider. Lastly, we report the results our replications, highlighting the importance of wide-scale replication efforts.

@@

\specialcell{Citation\\(Psychological Science; 2015)} & Expt. & \specialcell{Open\\Data?} & \specialcell{Open\\Materials?} & \specialcell{On\\MTurk?} & \specialcell{N\\(orig)} & \specialcell{Power\\Standard} & \specialcell{N\\(rep)} & \specialcell{Instructor\\Fidelity}\\ 
  \hline
Atir, Rosenzweig, \& Dunning & 1b & No & Yes & Yes & 202 & Other &  50 & 6.67 & \\ 
Ko, Sadler, \& Galinksy & 2 & Yes & Some & No &  40 & Original &  40 & 4.67 &\\ 
Lewis \& Oyserman & 4 & No & Yes & Yes & 122 & 80\% power & 128 & 6.67 &\\ 
Liverence \& Scholl & 1 & No & Some & No &  18 & Original* &  19 & 5.33 &\\ 
Proudfoot, Kay, \& Koval & 1 & No & No & Yes &  80 & 80\% power &  84 & 6.67&\\ 
Scopelliti, Loewenstein, \& Vosgerau & 3 & Yes & Yes & Yes & 550 & Other & 124 & 5.67& \\ 
Sofer et al. & 1 & Yes & Yes & No & 48 & Other &  95 & 4.33& \\ 
Storm \& Stone & 3 & No & No & No & 48 & Original &  61 & 4.00& \\ 
Wang et al. & 2 & No & No & No & 219 & 80\% power & 397 & 4.33&\\ 
Xu \& Franconeri & 1a & No & No & No &  12 & Other* &  27 & 5.00& \\ 
Zaval, Markowitz, \& Weber & 1 & Yes & Yes & Yes & 312 & Original & 321 & 5.00&\\ 
   \hline
\end{tabular}
\caption{\label{tab:summary} Summary characteristics of original studies and our replications. All project materials available publicly at osf.io/98ta4/files/, and links to individual project preregistrations, reports, and web experiments are available at osf.io/98ta4/wiki/. * marks projects where the number of trials was modified.} 
\label{tab:summary}
\end{table}

@@

RXDH, ENS, and MCF designed the project, supported data planning and data collection for all projects, analyzed the data, and wrote the paper. CA, JMA, RC, EH, MK, AL, SR, JR, SS, JS, and JT planned individual projects, programmed studies, collected data, analyzed data, and gave feedback on the paper.

@@

Though others' courses that incorporate replication projects will look quite different from our own, we recommend having the replication as a centerpiece of the course and giving ample opportunity for feedback throughout the study design. Without a central focus on the replication project, students will not have the time or motivation to complete a high quality project; without multiple opportunities for extensive (and often interactive) instructor feedback, the quality of the final projects will decline.

@@

In our class, students selected projects based on interest, rather than via a systematic sampling strategy. After selecting a paper, students chose a particular study from within that paper based on their judgment of interest and feasibility. We encouraged students with little programming background to opt for methodologically simple studies, and more advanced students to take on a challenging design. This level of freedom is not ideal for less advanced courses, and the set of available replication studies should likely be curated for students at lower levels. For example, in a large undergraduate class it may be logistically simpler for the instructors to choose a small set of replication projects that students can select from. This step would both give instructors a greater degree of familiarity with the literature and allow for the development of more scaffolding with respect to the development of experimental materials. A mid-level course might split the difference between these two models, providing students with a broader -- but still curated -- list of potential projects.

@@

Some of these tools can feel inaccessible or intimidating to students with a more limited programming background. Indeed, our students, who came from a variety of disciplinary backgrounds, typically have relatively little experience with web-based empirical studies or literate, reproducible data analysis. Yet by the end of the course, all gain sufficient proficiency with the suite of technical and conceptual tools necessary to carry out an independent project (with support from their peers and the course staff). Our experience suggests that it is possible to convey key concepts in sufficient depth that students can learn to use a broad range of open and reproducible tools for their projects, even if mastery will require further experience.

@@

Determining appropriate sample sizes is a major challenge in replication research [@button2013;@simonsohn2015]. Sample planning based on analysis of the achieved effect size in a previously-published study is problematic because of the likelihood of inflation of published effect sizes due to the "winner's curse" [@hoenig2001;@button2013]. But -- especially within the constraints of a limited budget -- it can be impossible to follow more conservative guidelines in all cases. For example, @simonsohn2015 recommends 2.5x the original sample, which can be feasible for small or under-powered original studies but may lead to impractical or unnecessary recommendations for studies that were initially large and/or adequately powered. In general, samples for course projects will be limited by the available resources. For all replication research of the type advocated here, achieving high power for a smaller number of studies, perhaps conducted by groups, is preferable to conducting more lower-power studies, because results are much more likely to make a contribution to the replication literature (as well as to be interpretable by students). Regardless of the eventual sample collected, we find that the discussion of sample size determination is one of the most eye-opening parts of our class and so we encourage instructors to allow students to participate in this planning process.

@@

Over half of our students contacted original authors to request study materials or clarification on methods and/or analysis, but for the purposes of our class, we chose not to require contact with all original authors. We strongly believe methods and materials should be open to allow for replication and productive science [@nosek2015], and hoped to empower students to engage with and build on the published literature directly -- without personal relationships or even personal contact. That is, we chose to do good-faith replications given all information in the original articles and supplemental online materials. Authors were contacted only in cases where there were issues about access to materials or ambiguities in design or analysis. There are benefits of contacting original authors, however. In addition to professional courtesy, authors can provide valuable feedback or guidance (albeit at the cost of some imposition) prior to data collection, potentially heading off post-hoc debates about methodological choices. In the end, it is up to individual instructors to decide what they believe is best for their students in this area.

@@

Ethical review standards vary across countries and are idiosyncratic across institutions, but we suspect certain ethical concerns will commonly arise when attempting to implement class replication projects. First, instructors may first be concerned that ethics approval is not possible within the time constraints of the class. To mitigate this issue, we suggest contacting your ethics body well in advance of the class to determine whether to create a standard protocol that encompasses all replications that will be conducted (as we did), or to determine a timeline for submitting individual protocols very early in the course. Both types of protocol will be simplified if the study or pool of studies to be replicated is pre-selected by instructors. Second, review boards may be concerned about risks particular to the collection of data by students, e.g., lack of debriefing or greater risk of breach of confidentiality due to novice experimenters. We recommend building training around ethical issues into the syllabus of replication-based classes. This practice is both positive for students and mitigates potential ethics approval concerns via the inclusion of explicit training.
